---
title: "Growth Curve Analysis 2"
author: "Masoud Jasbi"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r packages, include=FALSE}
library(tidyverse)
library(brms)
library(ggeffects)
library(QurvE)
library(gridExtra)
```

# Read Data

Load data from the results folder for the AL and Random methods:

```{r LoadData}
new_results <- 
  read_delim("../results/finalized_1000_pos_results_new.txt", delim = " ") %>%
  filter(Seed == "average")
```

Limiting the analysis to only F1 scores and removing the Treebanks with no results.

```{r}
al_500_pos_results <-
  new_results %>%
  filter(Metric=="f1", Method=="al", Treebank!="UD_German-HDT", Language!="Latin", Treebank!="UD_Manx-Cadhan")

random_500_pos_results <-
  new_results %>%
  filter(Metric=="f1", Method=="random", Treebank!="UD_German-HDT", 
         Treebank!="UD_Latin-ITTB", Treebank!="UD_Latin-Perseus", , Treebank!="UD_Manx-Cadhan",
         Treebank!="UD_Latin-UDante", Treebank!="UD_Latin-PROIEL", Treebank!="UD_Latin-LLCT")
```

# Statistical Models

Parameters: 

the upper asymptote $a$, 
the growth rate $r$, 
the inflection point $i$ along the x-axis
Displacement along the y-axis $b$
Displacement along the x-axis $i$
$e$ is Euler's number): 

$Y_{L} = a-(a-b)\times e^{-(cx)}$

# Fitting All the Models

Creating two empty data frames for storing model predictions and curves + derivatives `r Treebank_GrowthRates` and another empty data frame for model parameteres `r GrowthCurve_Models`.

```{r}
Treebank_GrowthRates <- data.frame()
GrowthCurve_Models <- data.frame()
```

```{r}
Data_Set <- random_500_pos_results
Treebank_List <-unique(Data_Set$Treebank)
```

The loop below takes the data fro each Treebank, fits the growth curve model, stores the model parameters, stores the model predicitons as well as the first derivative in the empty data frames created above.

```{r}
for (Treebank_Label in Treebank_List){
  
Treebank_Data <-
  Data_Set %>% 
  filter(Treebank==Treebank_Label)

Treebank_Model <- brm(
  bf(Value  ~ upperAsymptote - (upperAsymptote -lowerAsymptote) * exp(-(growthRate*Select_size)^delta),
     upperAsymptote ~ 1,
     lowerAsymptote ~1,
     growthRate ~ 1,
     delta ~ 1,
     nl = TRUE),
  data = Treebank_Data,
  prior = c(
    prior(uniform(0, 1), nlpar = "upperAsymptote", lb = 0, ub = 1),
    prior(uniform(0, 1), nlpar = "lowerAsymptote", lb = 0, ub = 1),
    prior(uniform(0, 10), nlpar = "growthRate", lb = 0, ub = 10),
    prior(uniform(0, 10), nlpar = "delta", lb = 0, ub = 10)
  ),
  file = paste("sample500random/", Treebank_Label, "_random", sep=""),
  iter = 4000)

# Model Parameters Dataframe
Treebank_Model_fixefs <- as.data.frame(fixef(Treebank_Model))
Treebank_Model_fixefs$Treebank <- Treebank_Label
GrowthCurve_Models <- rbind(GrowthCurve_Models, Treebank_Model_fixefs)

# First Derivative Curves
Treebank_Model_predictions <- as.data.frame(ggpredict(Treebank_Model))

Treebank_Model_predictions$upperAsymptote <- fixef(Treebank_Model)[1]
Treebank_Model_predictions$lowerAsymptote <- fixef(Treebank_Model)[2]
Treebank_Model_predictions$growthRate <- fixef(Treebank_Model)[3]
Treebank_Model_predictions$delta <- fixef(Treebank_Model)[4]

Treebank_GrowthRate <-
  Treebank_Model_predictions %>%
  filter(Select_size.x!=0) %>%
  mutate(Growth_Rate = (upperAsymptote - lowerAsymptote)*growthRate^delta*delta*Select_size.x^(delta-1)*exp(-(growthRate*Select_size.x)^delta))

Treebank_GrowthRate$Treebank <- Treebank_Label

Treebank_GrowthRates <- rbind(Treebank_GrowthRates, Treebank_GrowthRate)
}
```

Now saving the dataframes for the AL model results and predictions.

```{r}
write.csv(Treebank_GrowthRates, "New500_GrowthRates_AL.csv")
write.csv(GrowthCurve_Models, "New500_GrowthCurveModels_AL.csv")
```


```{r}
write.csv(Treebank_GrowthRates, "New500_GrowthRates_random.csv")
write.csv(GrowthCurve_Models, "New500_GrowthCurveModels_random.csv")
```

# Russian Treebanks

```{r}
Treebank_List <- list.files(path = "Russian500", pattern = "*.rds")
```

```{r}
Russian_Models <- data.frame()
Russian_Curves <- data.frame()

for (Treebank_Label in Treebank_List){
  
Treebank_Model <- readRDS(paste("Russian500/", Treebank_Label, sep=""))

# Model Parameters Dataframe
Treebank_Model_fixefs <- as.data.frame(fixef(Treebank_Model))
Treebank_Model_fixefs$Treebank <- Treebank_Label
Russian_Models <- rbind(Russian_Models, Treebank_Model_fixefs)

# First Derivative Curves
Treebank_Model_predictions <- as.data.frame(ggpredict(Treebank_Model))

Treebank_Model_predictions$upperAsymptote <- fixef(Treebank_Model)[1]
Treebank_Model_predictions$lowerAsymptote <- fixef(Treebank_Model)[2]
Treebank_Model_predictions$growthRate <- fixef(Treebank_Model)[3]
Treebank_Model_predictions$delta <- fixef(Treebank_Model)[4]

Treebank_Model_predictions <-
  Treebank_Model_predictions %>%
  filter(Select_size.x!=0) %>%
  mutate(Growth_Rate = (upperAsymptote - lowerAsymptote)*growthRate^delta*delta*Select_size.x^(delta-1)*exp(-(growthRate*Select_size.x)^delta))

Treebank_Model_predictions$Treebank <- Treebank_Label

Russian_Curves <- rbind(Russian_Curves, Treebank_Model_predictions)
}
```

```{r}
Russian_Models <-
  Russian_Models %>%
  rownames_to_column("Parameter") %>%
  separate(Parameter, c("Parameter","Whatever"), "_") %>%
  select(-Whatever) %>%
  separate(Treebank, c("Dependency", "Language", "Treebank", "Method", "extension"), "[-_.]")
```

```{r}
Russian_Curves <-
  Russian_Curves %>%
  separate(Treebank, c("Dependency", "Language", "Treebank", "Method", "extension"), "[-_.]")
```

## Growth Curves

```{r}
Russian_Curves %>%
  ggplot(aes(Select_size.x, Select_size.predicted, color=Method)) +
  geom_line()+
  geom_ribbon(aes(ymin = Select_size.conf.low, ymax = Select_size.conf.high), alpha=0.3, linetype=0) + 
  theme_linedraw() + facet_wrap("Treebank") + 
  labs(
    x = "Training Set Size",
    y = "F1 Scores")
```

```{r}
Russian_Models$Parameter <- recode_factor(Russian_Models$Parameter, 
                                          upperAsymptote = "Upper Asymptote", growthRate = "Growth Rate")

Russian_Models %>%
  filter(Parameter!="delta", Parameter!="lowerAsymptote") %>%
  ggplot(aes(Treebank,Estimate, color=Method)) + 
  geom_point(size=2) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap("Parameter", scales = "free_y") + theme_linedraw() +
  theme(axis.text.x = element_text(angle = 75, hjust = 1), legend.position = "inside", legend.position.inside = c(0.9,0.8), text = element_text(size = 16))
```

```{r}
Russian_Models %>%
  filter(Parameter == "Growth Rate" | Parameter == "Upper Asymptote")
```

# Irish Treebanks

```{r}
Treebank_List <- list.files(path = "Irish", pattern = "*.rds")
```

```{r}
Irish_Models <- data.frame()
Irish_Curves <- data.frame()

for (Treebank_Label in Treebank_List){
  
Treebank_Model <- readRDS(paste("Irish/", Treebank_Label, sep=""))

# Model Parameters Dataframe
Treebank_Model_fixefs <- as.data.frame(fixef(Treebank_Model))
Treebank_Model_fixefs$Treebank <- Treebank_Label
Irish_Models <- rbind(Irish_Models, Treebank_Model_fixefs)

# First Derivative Curves
Treebank_Model_predictions <- as.data.frame(ggpredict(Treebank_Model))

Treebank_Model_predictions$upperAsymptote <- fixef(Treebank_Model)[1]
Treebank_Model_predictions$lowerAsymptote <- fixef(Treebank_Model)[2]
Treebank_Model_predictions$growthRate <- fixef(Treebank_Model)[3]
Treebank_Model_predictions$delta <- fixef(Treebank_Model)[4]

Treebank_Model_predictions <-
  Treebank_Model_predictions %>%
  filter(Select_size.x!=0) %>%
  mutate(Growth_Rate = (upperAsymptote - lowerAsymptote)*growthRate^delta*delta*Select_size.x^(delta-1)*exp(-(growthRate*Select_size.x)^delta))

Treebank_Model_predictions$Treebank <- Treebank_Label

Irish_Curves <- rbind(Irish_Curves, Treebank_Model_predictions)
}
```

```{r}
Irish_Models <-
  Irish_Models %>%
  rownames_to_column("Parameter") %>%
  separate(Parameter, c("Parameter","Whatever"), "_") %>%
  select(-Whatever) %>%
  separate(Treebank, c("Dependency", "Language", "Treebank", "Method", "extension"), "[-_.]")
```

```{r}
Irish_Curves <-
  Irish_Curves %>%
  separate(Treebank, c("Dependency", "Language", "Treebank", "Method", "extension"), "[-_.]")
```

## Growth Curves

```{r}
Irish_CurvePlot <-
  Irish_Curves %>%
  filter(Treebank=="IDT")%>%
  ggplot(aes(Select_size.x, Select_size.predicted, color=Method)) +
  geom_line()+
  geom_ribbon(aes(ymin = Select_size.conf.low, ymax = Select_size.conf.high), alpha=0.3, linetype=0) + 
  theme_linedraw() + facet_wrap("Treebank") + 
  labs(
    x = "Training Set Size",
    y = "F1 Scores") + theme(legend.position = "none")
```

```{r}
Irish_Models$Parameter <- recode_factor(Irish_Models$Parameter, 
                                          upperAsymptote = "Upper Asymptote", growthRate = "Growth Rate")

Irish_GrowthratePlot <-
  Irish_Models %>%
  filter(Treebank=="IDT")%>%
  filter(Parameter!="delta", Parameter!="lowerAsymptote", Parameter!="Upper Asymptote") %>%
  ggplot(aes(Treebank,Estimate, color=Method)) + 
  geom_point(size=2) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap("Parameter", scales = "free_y") + theme_linedraw() +
  theme(axis.text.x = element_text(angle = 75, hjust = 1), text = element_text(size = 16)) +
  labs(
    x = "",
    y = "Estimated Growth Rate")

Irish_UpperAsymptotePlot <-
  Irish_Models %>%
  filter(Treebank=="IDT")%>%
  filter(Parameter!="delta", Parameter!="lowerAsymptote", Parameter!="Growth Rate") %>%
  ggplot(aes(Treebank,Estimate, color=Method)) + 
  geom_point(size=2) +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap("Parameter", scales = "free_y") + theme_linedraw() +
  theme(axis.text.x = element_text(angle = 75, hjust = 1), legend.position = "inside", legend.position.inside = c(0.9,0.8), text = element_text(size = 16)) + theme(legend.position = "none") +
  labs(
    x = "",
    y = "F1 Score")
```

```{r}
grid.arrange(Irish_UpperAsymptotePlot, Irish_GrowthratePlot, layout_matrix = matrix(c(1,1,2,2,2), nrow = 1))
```


# Reloading All Models

## Random Method

```{r}
Treebank_List_Random <- list.files(path = "sample500random", pattern = "*.rds")

All_Models_Random <- data.frame()
All_Curves_Random <- data.frame()
```

```{r}
for (Treebank_Label in Treebank_List_Random){
  
Treebank_Model <- readRDS(paste("sample500random/", Treebank_Label, sep=""))

# Model Parameters Dataframe
Treebank_Model_fixefs <- as.data.frame(fixef(Treebank_Model))
Treebank_Model_fixefs$Treebank <- Treebank_Label
All_Models_Random <- rbind(All_Models_Random, Treebank_Model_fixefs)

# First Derivative Curves
Treebank_Model_predictions <- as.data.frame(ggpredict(Treebank_Model))

Treebank_Model_predictions$upperAsymptote <- fixef(Treebank_Model)[1]
Treebank_Model_predictions$lowerAsymptote <- fixef(Treebank_Model)[2]
Treebank_Model_predictions$growthRate <- fixef(Treebank_Model)[3]
Treebank_Model_predictions$delta <- fixef(Treebank_Model)[4]

Treebank_Model_predictions <-
  Treebank_Model_predictions %>%
  filter(Select_size.x!=0) %>%
  mutate(Growth_Rate = (upperAsymptote - lowerAsymptote)*growthRate^delta*delta*Select_size.x^(delta-1)*exp(-(growthRate*Select_size.x)^delta))

Treebank_Model_predictions$Treebank <- Treebank_Label

All_Curves_Random <- rbind(All_Curves_Random, Treebank_Model_predictions)
}

All_Models_Random$Method <- "Random"
All_Curves_Random$Method <- "Random"
```

```{r}
write.csv(All_Models_Random, "All500_Models_Random")
write.csv(All_Curves_Random, "All500_Curves_Random")
```

## AL Method

```{r}
Treebank_List_AL <- list.files(path = "sample500AL", pattern = "*.rds")

All_Models_AL <- data.frame()
All_Curves_AL <- data.frame()
```

```{r}
for (Treebank_Label in Treebank_List_AL){
  
Treebank_Model <- readRDS(paste("sample500AL/", Treebank_Label, sep=""))

# Model Parameters Dataframe
Treebank_Model_fixefs <- as.data.frame(fixef(Treebank_Model))
Treebank_Model_fixefs$Treebank <- Treebank_Label
All_Models_AL <- rbind(All_Models_AL, Treebank_Model_fixefs)

# First Derivative Curves
Treebank_Model_predictions <- as.data.frame(ggpredict(Treebank_Model))

Treebank_Model_predictions$upperAsymptote <- fixef(Treebank_Model)[1]
Treebank_Model_predictions$lowerAsymptote <- fixef(Treebank_Model)[2]
Treebank_Model_predictions$growthRate <- fixef(Treebank_Model)[3]
Treebank_Model_predictions$delta <- fixef(Treebank_Model)[4]

Treebank_Model_predictions <-
  Treebank_Model_predictions %>%
  filter(Select_size.x!=0) %>%
  mutate(Growth_Rate = (upperAsymptote - lowerAsymptote)*growthRate^delta*delta*Select_size.x^(delta-1)*exp(-(growthRate*Select_size.x)^delta))

Treebank_Model_predictions$Treebank <- Treebank_Label

All_Curves_AL <- rbind(All_Curves_AL, Treebank_Model_predictions)
}

All_Models_AL$Method <- "AL"
All_Curves_AL$Method <- "AL"
```

```{r}
write.csv(All_Models_AL, "All500_Models_AL")
write.csv(All_Curves_AL, "All500_Curves_AL")
```

```{r}
All_Models <- rbind(All_Models_AL, All_Models_Random)
All_Curves <- rbind(All_Curves_AL, All_Curves_Random)
```

```{r}
write.csv(All_Models, "All500_Models")
write.csv(All_Curves, "All500_Curves")
```

## Cleanup

```{r}
All_Models <- read.csv("All500_Models")
All_Curves <- read.csv("All500_Curves")
```


```{r}
All_Models <-
  All_Models %>%
  rename(Parameter="X") %>%
  separate(Parameter, c("Parameter","Whatever"), "_") %>%
  select(-Whatever) %>%
  separate(Treebank, c("Dependency", "Treebank", "model_extension"), "_")
#All_Models$Treebank <-str_remove(All_Models$Treebank, "_model.rds")
#All_Models$Treebank <-str_remove(All_Models$Treebank, "UD_")
```

```{r}
All_Curves <-
  All_Curves %>%
  separate(Treebank, c("Dependency", "Treebank", "model_extension"), "_")
```

## Growth Curves

```{r}
All_Curves %>%
  ggplot(aes(Select_size.x, Select_size.predicted, color=Method)) +
  geom_line()+
  geom_ribbon(aes(ymin = Select_size.conf.low, ymax = Select_size.conf.high), alpha=0.3, linetype=0) + 
  theme_linedraw() + facet_wrap(.~Treebank)
```

```{r}
All_Models$Parameter <- recode_factor(All_Models$Parameter, 
                                          upperAsymptote = "Upper Asymptote", growthRate = "Growth Rate")

All_Models %>%
  filter(Parameter=="Growth Rate",
         Treebank!="Telugu-MTG", Treebank!="Chinese-GSDSimp", Treebank!="Lithuanian-HSE", Estimate<0.1) %>%
  ggplot(aes(Treebank,Estimate, color=Method)) + 
  geom_point() +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap("Parameter", scales = "free_y") + theme_linedraw() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1), legend.position = "inside", legend.position.inside = c(0.95,0.8)) 

All_Models %>%
  filter(Parameter=="Growth Rate",
         Treebank!="Telugu-MTG", Treebank!="Chinese-GSDSimp", Treebank!="Lithuanian-HSE", Estimate>0.1)
```

```{r}
All_Models %>%
  filter(Parameter=="Upper Asymptote", Treebank!="Telugu-MTG", Treebank!="Chinese-GSDSimp", Treebank!="Lithuanian-HSE", Estimate>0.8) %>%
  ggplot(aes(Treebank,Estimate, color=Method)) + 
  geom_point() +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap("Parameter", scales = "free_y") + theme_linedraw() +
  theme(axis.text.x = element_text(angle = 75, hjust = 1), legend.position = "inside", legend.position.inside = c(0.08,0.2)) 
```

```{r}
All_Models %>%
  filter(Parameter=="Upper Asymptote", Treebank!="Telugu-MTG", Treebank!="Chinese-GSDSimp", Treebank!="Lithuanian-HSE") %>%
  arrange(Estimate, decreasing=T)
```

```{r}
All_Models %>%
  filter(Parameter=="Growth Rate") %>%
  ggplot(aes(Parameter,Estimate, color=Method)) + 
  geom_point() +
  geom_linerange(aes(ymin = Q2.5, ymax = Q97.5)) +
  facet_wrap(Parameter~Treebank, scales = "free_y") + theme_linedraw()
```


```{r}
All_Models_Average <-  
  All_Models %>%
  filter(Parameter=="Upper Asymptote" | Parameter=="Growth Rate") %>%
  group_by(Method, Parameter) %>%
  summarise(Average_F1 = mean(Estimate), Average_Lower = mean(Q2.5), Average_Higher = mean(Q97.5))
```

```{r}
All_Models_Average %>%
  ggplot(aes(Method, Average_F1, color=Method)) +
  geom_point() +
  geom_linerange(aes(ymin=Average_Lower, ymax=Average_Higher)) +
  facet_wrap("Parameter", scales = "free_y") +
  theme_linedraw()
```

# Refitting the Models that Failed 

```{r}
Individual_Model <- brm(
  bf(Value  ~ upperAsymptote - (upperAsymptote -lowerAsymptote) * exp(-(growthRate*Select_size)^delta),
     upperAsymptote ~ 1,
     lowerAsymptote ~1,
     growthRate ~ 1,
     delta ~ 1,
     nl = TRUE),
  data = filter(al_500_pos_results,Treebank=="UD_Russian-Taiga"),
  prior = c(
    prior(uniform(0, 1), nlpar = "upperAsymptote", lb = 0, ub = 1),
    prior(uniform(0, 1), nlpar = "lowerAsymptote", lb = 0, ub = 1),
    prior(uniform(0, 10), nlpar = "growthRate", lb = 0, ub = 10),
    prior(uniform(0, 10), nlpar = "delta", lb = 0, ub = 10)
  ),
  file = "UD_Russian-Taiga_AL",
  iter = 30000)
```

